---
title: "Canopy C and N"
author: "Beni Stocker"
date: "2024-02-24"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(LambertW)
```

## Model description

CN-model contains the P-model for photosyntheis, which is formulated as
a big-leaf model, whereby CO$_2$ assimilation is to be understood as the
ecosystem gross primary production (GPP). Accordingly, also
$V_\mathrm{cmax}$ and $R_d$ are to be understood as quantities
representative for canopy-totals and in units of per ground area. From
the formulation of the P-model (Stocker et al., 2020), it follows that
$V_\mathrm{cmax}$ scales in proportion to the light absorbed by the
canopy ($I_\mathrm{abs}$). Using the Beer-Lambert law for light
absorption, a saturating relationship between (canopy total)
$V_\mathrm{cmax}$ and the leaf area index (LAI) follows.

$$
V_\mathrm{cmax} = \underbrace{\varphi_0 \frac{\chi + \kappa}{\chi + 2\gamma} I_0}_{c_0} \left(1 - e^{-kL} \right)
$$

The metabolic leaf N ($N_v$) is taken to be proportional to
$V_\mathrm{cmax}$ at standard temperature (25°C, $V_\mathrm{cmax25}$).

$$
\begin{aligned}
N_v &= b_v V_{cmax25} \\
    &= \underbrace{b_v f_t^{-1} c_0}_{c_0^\ast} \left(1 - e^{-kL} \right)
\end{aligned}
$$
At the leaf-level, the amount of metabolic N per unit leaf area is 

$$
N_v^l = N_v/L
$$

The structural leaf N ($N_s$) is modelled as a linear function of
$N_v^l$. 

$$
N_s^l = a_s + b_s N_v^l
$$

The parameter $a_s$ determines the minimum investment into structures as
$N_v^l$ tends toward zero when $L$ tends towards infinity
(theoretically).

The total leaf N is the sum of metabolic and structural N at the leaf
level ($N^l = N_v^l + N_s^l$). At the canopy-level, this is

$$
\begin{aligned}
N_\mathrm{fol} &= L (N_v^l + N_s^l) \\
               &= a_s L + (1 + b_s)N_v \\
               &= a_s L + (1 + b_s) c_0^\ast \left(1 - e^{-kL} \right)
\end{aligned}
$$

Leaf C per unit leaf area (which is largely equivalent to leaf mass per
area, LMA) is calculated by multiplying $N_s^l$ with a constant factor
that relates canopy structural N to structural C. 
$$
C^l = b_c\;N_s^l
$$

The canopy-total C in foliage is 
$$
\begin{aligned}
C_\mathrm{fol} &= C^l L \\
               &= L b_c N_s^l \\
               &= L b_c (a_s + b_s N_v^l) \\
               &= L b_c a_s + b_c b_s N_v \\ 
               &= b_c \left( a_s L + b_s c_0^\ast \left(1 - e^{-kL} \right) \right)
\end{aligned}
$$

This is a transcendental function of the form 
$$
\alpha(1-e^{-kL}) + \beta L - \gamma = 0
$$ with:

-   $\alpha = b_s c_0^\ast$
-   $\beta = a_s$
-   $\gamma = C_\mathrm{fol}/b_c$

To express $L$ as a function of $C_\mathrm{fol}$, the Lambert-W function
$W_n$ has to be used:

$$
L = \frac{\gamma - \alpha}{\beta} + \frac{1}{k} W_n \left( \frac{\alpha k}{\beta} \exp \left( (\alpha - \gamma)k/\beta \right)  \right)
$$

## Demonstration

Let's apply this model for visualising:

-   $N_v$, $N_\mathrm{fol}$, $C_\mathrm{fol}$, $(C:N)_\mathrm{fol}$, and
    LMA versus LAI.
-   LAI versus $C_\mathrm{fol}$

Define and obtain values representing a temperate forest with a LAI of
5.

### Model parameters

```{r}
par <- list(
  
  # relationship between Vcmax25 and metabolic leaf N (including N in other enzymes 
  # assumed to scale with Vcmax25)
  nv_vcmax25 = 0.02 * 13681.77,  # b_v
  
  # constant ratio of C to structural N
  r_ctostructn_leaf = 40, #1.3 * 45.84125,   # b_c
  
  # y-axis intersection in the relationship of non-metabolic versus metabolic N per leaf area
  ncw_min = 0.056, #0.08 * 1.116222,  # a_s
  
  # slope in the relationship of non-metabolic versus metabolic N per leaf area
  r_n_cw_v = 1.23223, #0,  # b_s
  
  # light extinction coefficient 
  kbeer = 0.5,  # k
  
  molmass_c = 12.0107
)
```

### State variables

The value for $c_0$ is simulated dynamically in response to the
environment. It is

```{r}
# in rsofun: actnv_unitfapar = (nv_vcmax25 * vcmax25_unitiabs * dppfd / dayl), gpp_cnmodel.mod.f90, l. 238
# here, corresponds to b_v * c_0 * f_T^-1 (nv_vcmax25 * c0)
par$actnv_unitfapar <- 0.1079643 # 1.36396807E-06  # typical value, run analysis/example_cnmodel_ch_oe1.R
```

Define variable substitutes

-   $\alpha = b_c a_s$
-   $\beta = b_s c_0^\ast$

```{r}
par$alpha <- par$r_ctostructn_leaf * par$ncw_min
par$beta <- par$r_n_cw_v * par$actnv_unitfapar # actnv_unitfapar = nv_vcmax25 * c0
```

## Functions

```{r}
calc_n_v <- function(lai, par){
  par$actnv_unitfapar * (1 - exp(-par$kbeer * lai))
}

calc_n_vp <- function(lai, par){
  (1 + par$r_n_cw_v) * par$actnv_unitfapar * (1 - exp(-par$kbeer * lai))
}

calc_n_fol <- function(lai, par){
  par$ncw_min * lai + (1 + par$r_n_cw_v) * par$actnv_unitfapar * (1 - exp(-par$kbeer * lai))
}

calc_c_fol <- function(lai, par){
  lai * par$r_ctostructn_leaf * par$ncw_min + par$r_ctostructn_leaf * par$r_n_cw_v * par$actnv_unitfapar * (1 - exp(-par$kbeer * lai))
}

get_lai <- function(cleaf, par){

  alpha <- par$actnv_unitfapar * par$r_n_cw_v
  beta  <- par$ncw_min
  gamma <- cleaf / ( par$molmass_c * par$r_ctostructn_leaf ) 

  arg_to_lambertw <- alpha * par$kbeer / beta * exp( (alpha - gamma) * par$kbeer / beta )

  lai <- 1.0 / (beta * par$kbeer ) * ( -alpha * par$kbeer + gamma * par$kbeer + beta * W( arg_to_lambertw ) )

}

get_lai_approx_to_zero <- function( cleaf, par ){
  par$r_ctostructn_leaf / ( par$ncw_min * par$molmass_c * par$r_ctostructn_leaf )
}

get_lai_approx_to_inf <- function( cleaf, par ){
  1.0 / par$ncw_min * ( cleaf / (par$r_ctostructn_leaf * par$molmass_c) - par$actnv_unitfapar * par$r_n_cw_v)
}
```

## Visualisations

### Versus LAI

```{r}
df <- tibble(lai = seq(0, 10, by = 0.2)) |> 
  mutate(n_v = calc_n_v(lai, par),
         n_vp = calc_n_vp(lai, par),
         n_fol = calc_n_fol(lai, par),
         c_fol = calc_c_fol(lai, par)
         ) |> 
  mutate(lma = c_fol / lai,
         cn_fol = c_fol/n_fol
         )

df |> 
  pivot_longer(cols = c(cn_fol, lma, c_fol, n_fol, n_v, n_vp),
               names_to = "var",
               values_to = "val") |> 
  ggplot(aes(lai, val)) +
  geom_line() +
  facet_wrap(~var, scales = "free_y", ncol = 2)
```

### Versus leaf C

```{r}
df <- tibble(cleaf = seq(0, 300, by = 10)) |> 
  mutate(lai = get_lai(cleaf, par),
         lai_inf = get_lai_approx_to_inf(cleaf, par),
         lai_zero = get_lai_approx_to_zero(cleaf, par))

df |> 
  ggplot() +
  geom_line(aes(cleaf, lai)) +
  geom_line(aes(cleaf, lai_inf), linetype = "dotted") +
  theme_classic() +
  ylim(0, NA) +
  labs(x = expression(paste("leaf C (gC m"^-2,")")), 
       y = expression(paste("LAI (m"^2," m"^-2,")"))
  )

# ggsave(here::here("fig/lai_vs_leafc.pdf"), width = 4, height = 3)

# df |> 
#   ggplot() +
#   geom_line(aes(lai, cleaf)) +
#   geom_line(aes(lai_inf, cleaf), linetype = "dotted") +
#   theme_classic() +
#   xlim(0, NA)
```
