---
title: "N fixation"
author: "Beni Stocker"
date: "2025-04-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Calculate N fixation efficiency (dNfix/dCex) as a function of 
soil temperature. The functional form is chosen to be equal to
the (fitted) relationship between soil temperature and nitro-
genase activity derived by Houlton et al. (2008), Nature. The 
maximum efficiency is 0.21 gN/gC, which corresponds to the 
inverse of the "minimum cost of N-fixation" of 4.8 gC/gN given 
Gutschik (1981). At higher and lower temperature, efficiency de-
clines to zero.

```{r}
calc_eff_nfix <- function(soiltemp, par, norm = 1.25){
  out_eff_fix <- 1.0 / par[["minimumcostfix"]] * norm * 
      exp( par[["a_param_fix"]] + 
           par[["b_param_fix"]] * soiltemp * ( 1.0 - 0.5 * soiltemp / par[["fixoptimum"]] ) )

	return( out_eff_fix )	
}

par <- list(
  minimumcostfix = 4.8,
  fixoptimum = 25.15,
  a_param_fix = -3.62,
  b_param_fix = 0.27
)
```

```{r}
ggplot() +
  geom_function(fun = calc_eff_nfix, args = list(par = par), xlim = c(0, 50)) + 
  labs(
    x = "Soil temperature (°C)",
    y = "N fixation efficiency (gN/gC)"
  )
```

